### 2010-04-13 trunk/deliverables/D2_2_Ontology/fast_d2_2_3_ontology.tex
post deliverable submission start on third version
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@779 d8b0c59b-6544-0410-94d9-83a5658a058e

\begin{document}
% explicit hyphenations
\hyphenation{RDF-Re-po-si-to-ry}
\hyphenation{name-space}
\hyphenation{So-cket-A-dap-ter}

@@

\lohead{\includegraphics[width=4cm]{images/FAST_logo_transparent}}
%\cohead{\small\textcolor{fast@lightgrey}{\deliverableTitle}}
%\rohead{\small{\today}}
%\lofoot{\small\textcolor{fast@lightgrey}{Task Force Ontologies}}
\cofoot{\small{FAST $\bullet$ 216048 $\bullet$ \deliverableTitleShort{} $\bullet$ Page \thepage\ of \pageref{LastPage}}}
% \rofoot{\small{\thepage}}

@@

\begin{tabular}
	{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{1.4cm}|p{3.28cm}|p{1.6cm}|p{3.29cm}|}
	\hline
	\textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {28/02/2010} & {\textbf{Submitted}} & {27/02/2010}\\ \hline
	\textcolor{white}{\textbf{Status}} & \multicolumn{2}{l|}{} & \multicolumn{2}{l|}{final}\\ \hline
	\textcolor{white}{\textbf{Dissemination Level}} & \multicolumn{4}{l|}{Public $\boxtimes$ / Consortium $\square$}\\ \hline
	\textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{\deliverableNumber{} is the second iteration of the FAST Gadget ontology. The ontology provides a formal description of the conceptual model of FAST in general, and of the screen/screenflow architecture in particular. The document a number of successive intermediate representations which document the development process, and the final product in the form of an OWL ontology. Numerous code examples and visual representations of the ontology terms are provided.}\\ \hline
	\textcolor{white}{\textbf{Authors}} & \multicolumn{4}{p{10.85cm}|}{\authorOne, \authorTwo, \authorThree, \authorFour}\\
%	{} & \multicolumn{4}{l|}{}\\ 
%	{} & \multicolumn{4}{l|}{}\\ 
%	{} & \multicolumn{4}{l|}{}\\
	\hline
	\textcolor{white}{\textbf{Deliverable Owner}} & \multicolumn{2}{l|}{\authorOne} & \textbf{email} & {knud.moeller@deri.org} \\ \cline{4-5}
	\textcolor{white}{\textbf{(Partner)}} & \multicolumn{2}{l|}{} & \textbf{phone} & {+353 91 495086} \\ \hline
	\textcolor{white}{\textbf{Keywords}} & \multicolumn{4}{p{10.85cm}|}{FAST, conceptual model, ontology, OWL, RDFS}\\ \hline
\end{tabular}
\end{small}
\newpage

@@

In FAST, we have adopted the term \emph{gadget} for the small, Web-based software components which the project focusses on. There are small differences in meaning between this term and the term \emph{widget}, which is also widely used. However, for the purpose of this document, it should be noted that whenever we use the term \emph{gadget}, we also refer to what \emph{widget} at the same time.

@@

\begin{itemize}
	\item \textbf{Abstract Class Level}: At this level, resources are defined on a categorical level, differentiating between the various kinds of things that are relevant in FAST, as fundamental classes such as screens, forms or resources (see Sect.~\ref{sub:conceptualisation}). These are implemented technically as classes in OWL.
	\item \textbf{Prototypical Component Definition}: When a user of the FAST GVS is building a new screenflow or screen, they do so by combining existing building blocks and configuring them. These building blocks are not the fundamental classes of FAST, but instead they are specific instances of those classes. In other words, users do not combine the idea of a screen with the idea of a service, but instead they combine the specific ``Log In Screen'' with the ``Amazon Web Service''. Technically, these concrete building blocks are implemented as instances of the fundamental classes, specified by asserting particular settings such as label or icons, pre- and post-conditions, etc.
	\item \textbf{Instantiation within a Screenflow at Design-time}: Once a user selects a specific building block in the GVS and adds it to their new screen or screenflow project, they are in principle again instantiating from a conceptual to a concrete level: from the ``Log In Screen'' as such to a particular log in screen as used in that particular screenflow. Also in this stage, resources are implemented technically as instances.
	\item \textbf{Runtime}: At the final level of instantiation, the screenflow has been packaged as a gadget, deployed to a gadget platform and is now being used by an end-user. At this level, all gadget components exist only in the form of Web standards compliant code (HTML, JavaScript, CSS). For performance reasons, the FAST catalogue transforms any metadata still necessary, as well as definitions for screen pre- and post-conditions, from their original RDF representations into the JavaScript Object Notation (JSON)\footnote{\url{http://www.json.org/}}. At this level, the resources that make up the gadget have moved out of the scope of the FAST platform and therefore out of the scope of this deliverable.
\end{itemize}

@@

Neither OWL nor RDFS semantics offer functionality to represent a relationship such as the copy of a resource formally. For this reason, we introduce the concept of \emph{templates for RDF instances} in FAST. One resource (e.g., $LogInScreen$) will play the role of a \emph{template}, while the other resource (e.g., $LogInScreen_x$) will be called a \emph{copy}. Conceptually, copying a resource involves the following steps:
\begin{inparaenum}[(i)]
	\item Defining a named graph $G_1$ that contains all statements about this resource considered to be relevant.
	\item Creating a second graph $G_2$ with copies of all statements in $G_1$.
	\item During this copying process, the identifiers for the main resource and certain other resources in the graph will be changed from $I_{template}$ to $I_{copy}$.
\end{inparaenum}	
In practice, this procedure can be implemented using a SPARQL CONSTRUCT query, where the WHERE clause defines the graph pattern of the template, and the CONSTRUCT query will build the copy from it. The corresponding data on all three levels is shown in List.~\ref{list:templating}, using TriG syntax~\cite{bizer2004trig} (an extension of Turtle syntax for named graphs).

>>>>

\begin{document}
% explicit hyphenations
% \hyphenation{RDF-Re-po-si-to-ry}
% \hyphenation{name-space}
% \hyphenation{So-cket-A-dap-ter}

@@

\lohead{\includegraphics[width=4cm]{images/FAST_logo_transparent}}
%\cohead{\small\textcolor{fast@lightgrey}{\deliverableTitle}}
%\rohead{\small{\today}}
%\lofoot{\small\textcolor{fast@lightgrey}{Task Force Ontologies}}
%\rofoot{\small{\thepage}}
\cofoot{\small{FAST $\bullet$ 216048 $\bullet$ \deliverableTitleShort{} $\bullet$ Page \thepage\ of \pageref{LastPage}}}

@@

\begin{tabular}
	{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{1.4cm}|p{3.28cm}|p{1.6cm}|p{3.29cm}|}
	\hline
	\textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {28/02/2010} & {\textbf{Submitted}} & {27/02/2010}\\ \hline
	\textcolor{white}{\textbf{Status}} & \multicolumn{2}{l|}{} & \multicolumn{2}{l|}{final}\\ \hline
	\textcolor{white}{\textbf{Dissemination Level}} & \multicolumn{4}{l|}{Public $\boxtimes$ / Consortium $\square$}\\ \hline
	\textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{\deliverableNumber{} is the second iteration of the FAST Gadget ontology. The ontology provides a formal description of the conceptual model of FAST in general, and of the screen/screenflow architecture in particular. The deliverable contains a number of successive intermediate representations which document the development process, and the final product in the form of an OWL ontology. Numerous code examples and visual representations of the ontology terms are provided.}\\ \hline
	\textcolor{white}{\textbf{Authors}} & \multicolumn{4}{p{10.85cm}|}{\authorOne, \authorTwo, \authorThree, \authorFour}\\
%	{} & \multicolumn{4}{l|}{}\\ 
%	{} & \multicolumn{4}{l|}{}\\ 
%	{} & \multicolumn{4}{l|}{}\\
	\hline
	\textcolor{white}{\textbf{Deliverable Owner}} & \multicolumn{2}{l|}{\authorOne} & \textbf{email} & {knud.moeller@deri.org} \\ \cline{4-5}
	\textcolor{white}{\textbf{(Partner)}} & \multicolumn{2}{l|}{} & \textbf{phone} & {+353 91 495086} \\ \hline
	\textcolor{white}{\textbf{Keywords}} & \multicolumn{4}{p{10.85cm}|}{FAST, conceptual model, ontology, OWL, RDFS}\\ \hline
\end{tabular}
\end{small}
\newpage

@@

In FAST, we have adopted the term \emph{gadget} for the small, Web-based software components which the project focusses on. There are small differences in meaning between this term and the term \emph{widget}, which is also widely used. However, for the purpose of this document, it should be noted that whenever we use the term \emph{gadget}, we might also have used the \emph{widget} at the same time.

@@

\begin{itemize}
	\item \textbf{Abstract Class Level}: At this level, resources are defined on a categorical level, differentiating between the various kinds of things that are relevant in FAST, as fundamental classes such as screens, forms or resources (see Sect.~\ref{sub:conceptualisation}). These are implemented technically as classes in OWL.
	\item \textbf{Prototypical Component Definition}: When a user of the FAST GVS is building a new screenflow or screen, they do so by combining existing building blocks and configuring them. These building blocks are not the fundamental classes of FAST, but instead they are specific instances of those classes. In other words, users do not combine the idea of a screen with the idea of a service, but instead they combine the specific ``Log In Screen'' with the ``Amazon Web Service''. Technically, these concrete building blocks are implemented as instances of the fundamental classes, specified by asserting particular settings such as label or icons, pre- and post-conditions, etc.
	\item \textbf{Instantiation within a Screenflow at Design-time}: Once a user selects a specific building block in the GVS and adds it to their new screen or screenflow project, they are in principle again instantiating from a conceptual to a concrete level: from the ``Log In Screen'' as such to a particular log in screen as used in that particular screenflow. Also in this stage, resources are implemented technically as instances.
	\item \textbf{Runtime}: At the final level of instantiation, the screenflow has been packaged as a gadget, deployed to a gadget platform and is now being used by an end-user. At this level, all gadget components exist only in the form of Web standards compliant code (HTML, JavaScript, CSS). For performance reasons, the FAST catalogue transforms any metadata still necessary, as well as definitions for screen pre- and post-conditions, from their original RDF representations into the JavaScript Object Notation (JSON)\footnote{\url{http://www.json.org/}}. The JSON structures used at this level are described in \cite{palaghita2010catalogue_manual}. At this level, the resources that make up the gadget have moved out of the scope of the FAST platform and therefore out of the scope of this deliverable.
\end{itemize}

@@

Neither OWL nor RDFS semantics offer functionality to represent a relationship such as the copy of a resource formally. For this reason, we introduce the concept of \emph{templates for RDF instances} in FAST. Templates are similar to the notion of prototypes in programming languages such as JavaScript. One resource (e.g., $LogInScreen$) will play the role of a \emph{template}, while the other resource (e.g., $LogInScreen_x$) will be called a \emph{copy}. Conceptually, copying a resource involves the following steps:
\begin{inparaenum}[(i)]
	\item Defining a named graph $G_1$ that contains all statements about this resource considered to be relevant.
	\item Creating a second graph $G_2$ with copies of all statements in $G_1$.
	\item During this copying process, the identifiers for the main resource and certain other resources in the graph will be changed from $I_{template}$ to $I_{copy}$.
\end{inparaenum}	
In practice, this procedure can be implemented using a SPARQL CONSTRUCT query, where the WHERE clause defines the graph pattern of the template, and the CONSTRUCT query will build the copy from it. The corresponding data on all three levels is shown in List.~\ref{list:templating}, using TriG syntax~\cite{bizer2004trig} (an extension of Turtle syntax for named graphs).
